Network Configuration Synthesis with Abstract Topologies

### 2016-11-08 paper/abstraction/paper.tex
pass over intro

An important class of networks are the large \emph{structured networks}
commonly deployed in industrial data centers.  These networks run
many critical services; every second of downtime is
costly at best and dangerous at worst.  Unfortunately, it is an enormous
challenge to keep these complex networks up and
running 24/7~\cite{mahajan+:bgp-misconfiguration,feamster+:rcc,batfish,dc-failure-study}.
Moreover, while hardware faults, rodents, power failures and natural
disasters can all cause significant downtime, studies have shown that
\emph{human error}, which occurs during network configuration or
update, is the single largest cause of these
outages~\cite{juniper-study}.

% These networks often
% host a myriad of services and every second of downtime imposes significant
% financial cost or disruption to consumers.
Such networks typically have uniform, tree-shaped topologies that involve
hundreds or even thousands of devices.  Since it is not possible for humans
to think about configuring each individual device independently,
network operators typically classify the devices, assigning them a particular
\emph{role} such as ``top-of-rack router'' or ``tier-one aggregation router.''
Each such role is codified as a \emph{configuration template}---a macro
that may be instantiated with different concrete values to
generate a configuration for each device that plays the role.

Unfortunately, such templates have no particular semantics other than the
semantics
that one obtains from instantiating a template's parameters.
Moreover,
template construction is a low-level, manual process:  
network engineers do their
best to design templates despite the assembly-language-like primitives for network configuration.
They hope their templates
will work properly and test concrete instantiations of them using simulations, but
there is no support for verification of key network-wide properties 
such as connectivity and fault tolerance. In addition,
network engineers have no tools to help them design for network
evolution.  For instance, the process of adding new equipment to
expand the capacity of data center may not be compatible with the original
templates, or may cause a trickle-down of changes to be made network-wide.
Each change to a router configuration typically incurs
some risk of temporary instability in a network, so minimizing such changes
is an important design goal.  
%It is no wonder that a major
%network outage, such as the one that grounded all United Air flights
%in July 2015~\cite{x}, makes international news every few months.

To combat these problems, we present \sysname, a
new high-level language designed to
provide improved support for the configuration of large, structured networks.
\sysname allows network operators to define \emph{abstract network topologies}
that
classify devices in terms of their roles and how they are connected
to one another.  For instance, a programmer may
define separate roles for ``top-of-rack router'' and
``tier-one aggregation router'' and specify that every top-of-rack router
is connected to at least two different tier-one aggregation routers,
thereby providing a degree of fault tolerance if one aggregation router
were to fail.

To specify \emph{routing policy} for such abstract networks,
\sysname borrows notation from the recently-defined \propane
programming language~\cite{propane}.  As such, it is able to
specify the paths that traffic should flow along at a high-level of
abstraction.  However, whereas \propane policies
refer to paths of individual devices, \sysname policies
refer to paths of abstract roles.  The \sysname compiler compiles such
policies into templates, with one template per role, and these templates
may be instantiated to generate concrete configurations.

During compilation, the \sysname compiler also analyzes policy with respect to the
abstract topology and guarantees that the configurations will
faithfully implement the high-level policy on \emph{any} concrete
topology that satisfies the given connectivity invariants.  The compiler
will also optionally perform a second fault tolerance analysis that
determines the minimum number of link failures required to disconnect
\emph{any} valid concrete topology.

Finally, generated \sysname configurations are designed to
be \emph{evolution-friendly} in the sense that if an operator adds or
removes any single link or device, they are guaranteed that only
the configurations of adjacent devices need change (for instance, to
activate a port on which the new device is connected).  Doing so
helps minimize network instability as data centers are expanded or devices are
temporarily taken offline for maintenance.

As an aside, a pleasant side effect of this new compiler
infrastructure, designed to make life easier on network operators, is
that for structured networks, it scales substantially better than its
competitor \propane does: as the number of individual devices in a
network grows, the number of distinct roles will often stay constant.
Consequently, the \sysname compiler may be two orders of magnitude
faster than \propane for large structured networks.

@@

\para{Configuration synthesis with concrete topologies}
Many recent systems generate router configurations using high-level policy statements~\cite{narain:lisa05,narain+:configassure,dadc,propane} and the network's concrete topology.
%These systems eliminate entire classes of configuration errors by construction, and, in the case of
%Propane~\cite{propane}, guarantee that high-level policy is implemented correctly, even in the presence of
%arbitrary combinations of failures.
They are able to provide strong guarantees on configuration correctness but 
our conversations with operators at two major cloud providers reveal that they have two limitations that discourage adoption.  The first is abstraction mismatch. Operators of large networks think of the network, and develop management tools, using \emph{roles}, and not individual devices. They are naturally reluctant to use a system that does not
understand roles as a first-order abstraction. If they want to debug, analyze or
understand system output, they will have to consider
\emph{hundreds} of independent device configurations instead of just
\emph{a handful} of role configurations. Unfortunately, even if two devices play the same role,
there is no guarantee a concrete synthesis system will generate (syntactically) similar configurations.

%A second limitation is the inability to support network evolution. Networks evolve all the time as nodes and links are added or removed and the addresses of services are changed. A system that supports evolution would ensure that
%small changes in topology require small, local changes in network configurations as opposed to pervasive
%changes across many devices.  Engineers want to minimize changes to device configurations because each change runs the risk of poor, transient behavior due to poor router software or routing protocol dynamics~\cite{ratulbgpmisconfigs}.
%These networks must be highly
%available; it isn't plausible to bring down an entire data center network, or large portions of it, to make
%incremental changes.

Second, these systems become even more brittle in the face of network evolution.
Any change in network topology requires re-execution of a
concrete synthesis engine, from scratch,
on the changed topology.  The result may be
a completely different set of configurations.  Even if such configurations are
largely semantically similar to existing configurations, they may be syntactically different and cause considerable network disruption when updates are pushed to routers.

>>>>

%An important class of networks are the large \emph{structured networks}
%commonly deployed in industrial data centers.

Computer networks run many critical services, and every second of downtime is
costly at best and dangerous at worst.  Keeping these complex systems up and running 24/7 is an enormous challenge~\cite{mahajan+:bgp-misconfiguration,feamster+:rcc,batfish,dc-failure-study}.
%
While hardware faults, rodents, power failures and natural disasters all pose problems, studies have consistently shown that human error in network configuration is one of the most error-prone task and the leading cause of network outages~\cite{juniper-study,xx}.

To manage and configure large networks today, instead of considering individual devices, operators classify devices devices into {\em roles}. A role refers to specific functionality and is played by one or more devices. For instance, in a data center, roles may correspond to "top-of-rack," "aggregation," and "spine" switches; and in a backbone network, they may correspond to "core" and "border" routers. While the network may have hundreds or thousands of devices---a scale that is impossible for humans to handle---there tend to be a handful of roles.
%
Operators author a configuration {\em template} for each role. Templates are macros that can be instantiated with different concrete values, based on the network topology, to generate a configuration for each device.

Unfortunately, template construction is a low-level, manual process. Operators manually decompose network-wide routing and fault-tolerance policies (e.g., prefer one end-to-end path over another or no loss in connectivity after any one-link failure) across roles and encode them using assembly-language-like configuration primitives (e.g., role A should not transmit routing information from B to C). Such decomposition tasks are highly complex~\cite{propane}.

Worse, operators can never be sure that the generated templates correctly implement the policy,  beyond testing some concrete instantiations using simulation or verification tools~\cite{x,y}. Testing based on concrete instantiations is insufficient because large networks evolve continuously, as devices and links are added or taken offline for maintenance. Templates that work for the current network topology, may or may not work for future topologies. Updating the templates, if found to not work for a future topology, may cause trickle-down, network-wide configuration changes, which is highly disruptive. Building sophisticated tools for validating templates is difficult because the templates today have no particular semantics---which policies they implement and for which topologies---other than the semantics of the configurations that one obtains from instantiating them.

%It is no wonder that a major
%network outage, such as the one that grounded all United Air flights
%in July 2015~\cite{x}, makes international news every few months.

Given the challenges of generating and validating network configurations, one might think that operators will be receptive to systems that can generate provably-correct configurations from high-level policy specifications~\cite{x,y,z,configassure,propane}. However, our conversations with two major cloud providers reveal that operators of large networks are reluctant to use such systems. While they think of their network abstractly, in terms of roles, these systems require concrete topologies, with no understanding of roles as a first-order abstraction.
Even if two devices play the same role,
there is no guarantee that the systems will generate (syntactically) similar configurations; if the operators want to debug or analyze system output, they will have to consider
hundreds of device configurations instead of a handful of role configurations. These systems become even more brittle when the network evolves. Any change in network topology requires re-execution of the engine, from scratch, on the changed topology.  The result may be
a completely different set of configurations.

We present \sysname, a system to generate provably-correct role templates. It allows operators to input \emph{abstract network topologies} in terms of roles and their connectivity. For instance, they may specify roles for ``top-of-rack'' and ``aggregation" switches and specify that every top-of-rack switch connects to at least two aggregation switches (to tolerate the failure of a link to an aggregation switch). Other inputs to \sysname include the routing policy of the network and its fault-tolerance requirements (e.g., the number of link failures to tolerate). For routing policy, we borrow notation from Propane~\cite{propane},  which specifies traffic paths at a high-level of abstraction. But while Propane policies
refer concrete devices, \sysname policies refer to abstract roles.

Based on these inputs, the \sysname compiler generates one template per role. It guarantees that, when the templates are instantiated to configurations for {\em any} concrete topology that complies with the abstract topology, the result network will correctly implement the routing policy and meet the fault-tolerance requirements. We show how fault-tolerance analysis can be done over abstract topology itself using \todo{say something cool about about the inference rules and such}.

\sysname templates are also evolution-friendly. When the network migrates from one concrete topology to another, both of which comply with the abstract topology, it is guaranteed that only the configurations of devices that acquire or lose a neighboring device needs to change (e.g., to activate a port on which the new device is connected).  No other device in the network is impacted.

\todo{Talk about evaluation result here ... roll the point about faster compilation there.}
As an aside, a pleasant side effect of this new compiler
infrastructure, designed to make life easier on network operators, is
that for structured networks, it scales substantially better than its
competitor \propane does: as the number of individual devices in a
network grows, the number of distinct roles will often stay constant.
Consequently, the \sysname compiler may be two orders of magnitude
faster than \propane for large structured networks.

@@

%\para{Configuration synthesis with concrete topologies}
%Many recent systems generate router configurations using high-level policy statements~\cite{narain:lisa05,narain+:configassure,dadc,propane} and the network's concrete topology.
%%These systems eliminate entire classes of configuration errors by construction, and, in the case of
%%Propane~\cite{propane}, guarantee that high-level policy is implemented correctly, even in the presence of
%%arbitrary combinations of failures.
%They are able to provide strong guarantees on configuration correctness but
%our conversations with operators at two major cloud providers reveal that they have two limitations that discourage adoption.  The first is abstraction mismatch. Operators of large networks think of the network, and develop management tools, using \emph{roles}, and not individual devices. They are naturally reluctant to use a system that does not
%understand roles as a first-order abstraction. If they want to debug, analyze or
%understand system output, they will have to consider
%\emph{hundreds} of independent device configurations instead of just
%\emph{a handful} of role configurations. Unfortunately, even if two devices play the same role,
%there is no guarantee a concrete synthesis system will generate (syntactically) similar configurations.
%
%%A second limitation is the inability to support network evolution. Networks evolve all the time as nodes and links are added or removed and the addresses of services are changed. A system that supports evolution would ensure that
%%small changes in topology require small, local changes in network configurations as opposed to pervasive
%%changes across many devices.  Engineers want to minimize changes to device configurations because each change runs the risk of poor, transient behavior due to poor router software or routing protocol dynamics~\cite{ratulbgpmisconfigs}.
%%These networks must be highly
%%available; it isn't plausible to bring down an entire data center network, or large portions of it, to make
%%incremental changes.
%
%Second, these systems become even more brittle in the face of network evolution.
%Any change in network topology requires re-execution of a
%concrete synthesis engine, from scratch,
%on the changed topology.  The result may be
%a completely different set of configurations.  Even if such configurations are
%largely semantically similar to existing configurations, they may be syntactically different and cause considerable network disruption when updates are pushed to routers.
